---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
---

Placeholder file for the future R coding exercise.



## Getting Starting

install.packages('dslabs')
install.packages('tidyverse')
install.packages('ggplot2')




## Loading and Checking Data

```{r}
# Load dslabs package
library("dslabs")

# Look at help file for gapminder data
help(gapminder)

# Get an overview of data structure
str(gapminder)

# Get a summary of data
summary(gapminder)

# Determine the type of object gapminder is
class(gapminder)

```



## Processing Data

```{r}
# Load Tidyverse 
library(tidyverse)

# New object for African Countries
africadata <- gapminder %>%
  filter(continent == 'Africa')

# Check africadata
str(africadata)
summary(africadata)

# Create infant_mortality
inf_mortality <- africadata %>%
  select(infant_mortality, life_expectancy)

# Check inf_mortality
str(inf_mortality)
summary(inf_mortality)

# Create life_expectancy
lif_expect <- africadata %>%
  select(life_expectancy, population)

# Check lif_expect
str(lif_expect)
summary(lif_expect)

```



## Plotting

```{r}
# Install ggplot2
library(ggplot2)

# Plot life expectancy as a function of inf_mortality
ggplot(inf_mortality, aes(x = infant_mortality,
                          y = life_expectancy)) +
  geom_point() + labs(x = "Infant Mortality (per 1,000 births)",
                      y = "Life Expectancy (in years)")

# Plot life expectancy as a function of population size
ggplot(lif_expect, aes(x = log(population),
                          y = life_expectancy)) +
  geom_point() + labs(x = "Log Population", y = "Life Expectancy (in years)")

```



## More data processing

```{r}
# Find NAs
subset(africadata,is.na(inf_mortality))

with(subset(africadata, is.na(infant_mortality)), year)

# Filter year with most data
More <- africadata %>%
  filter(year == 2000)

# Check
summary(More)
str(More)

```



## More plotting

```{r}
# Plot life expectancy as a function of inf_mortality
ggplot(More, aes(x = infant_mortality,
                 y = life_expectancy)) +
  geom_point() + labs(x = "Infant Mortality (per 1,000 births)", 
                      y = "Life Expectany (in years)", 
                      title = "Infant Mortality and Life Expectancy in Africa 
                      (Year: 2000)")

# Plot life expectancy as a function of population size
ggplot(More, aes(x = log(population),
                          y = life_expectancy)) +
  geom_point() + labs(x = "Log Population", y = "Life Expectany (in years)", 
                      title = "Population and Life Expectancy in Africa 
                      (Year: 2000)")

```



## Fitting a simple model

```{r}
# lms
fit1 <- lm(life_expectancy ~ infant_mortality, data = More)
fit2 <- lm(life_expectancy ~ population, data = More)

# Results fit1
summary(fit1)

# Results fit2
summary(fit2)

```



## Conclusions

While the data does not show a statistically significant association between 
life expectancy and population size given our p-value is greater than our alpha 
level of 0.05 (p = 0.62), there was significant evidence of an association 
between life expectancy and infant mortality give out p-value is less than our 
alpha level of 0.05 (p < 0.001). Life expectancy decreases by 2.49 year on
average for every unit increase in infant mortality.

