---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
---
## Getting Starting

## Loading and Checking Data

```{r}
# Load dslabs package
library("dslabs")

# Look at help file for gapminder data
help(gapminder)

# Get an overview of data structure
str(gapminder)

# Get a summary of data
summary(gapminder)

# Determine the type of object gapminder is
class(gapminder)

```



## Processing Data

```{r}
# Load Tidyverse 
library(tidyverse)

# New object for African Countries
africadata <- gapminder %>%
  filter(continent == 'Africa')

# Check africadata
str(africadata)
summary(africadata)

# Create infant_mortality
inf_mortality <- africadata %>%
  select(infant_mortality, life_expectancy)

# Check inf_mortality
str(inf_mortality)
summary(inf_mortality)

# Create life_expectancy
lif_expect <- africadata %>%
  select(life_expectancy, population)

# Check lif_expect
str(lif_expect)
summary(lif_expect)

```



## Plotting

```{r}

# Plot life expectancy as a function of inf_mortality
ggplot(inf_mortality, aes(x = infant_mortality,
                          y = life_expectancy)) +
  geom_point() + labs(x = "Infant Mortality (per 1,000 births)",
                      y = "Life Expectancy (in years)")

# Plot life expectancy as a function of population size
ggplot(lif_expect, aes(x = log(population),
                          y = life_expectancy)) +
  geom_point() + labs(x = "Log Population", y = "Life Expectancy (in years)")

```



## More data processing

```{r}
# Find NAs
subset(africadata,is.na(inf_mortality))

with(subset(africadata, is.na(infant_mortality)), year)

# Filter year with most data
More <- africadata %>%
  filter(year == 2000)

# Check
summary(More)
str(More)

```



## More plotting

```{r}
# Plot life expectancy as a function of inf_mortality
ggplot(More, aes(x = infant_mortality,
                 y = life_expectancy)) +
  geom_point() + labs(x = "Infant Mortality (per 1,000 births)", 
                      y = "Life Expectany (in years)", 
                      title = "Infant Mortality and Life Expectancy in Africa 
                      (Year: 2000)")

# Plot life expectancy as a function of population size
ggplot(More, aes(x = log(population),
                          y = life_expectancy)) +
  geom_point() + labs(x = "Log Population", y = "Life Expectany (in years)", 
                      title = "Population and Life Expectancy in Africa 
                      (Year: 2000)")

```



## Fitting a simple model

```{r}
# lms
fit1 <- lm(life_expectancy ~ infant_mortality, data = More)
fit2 <- lm(life_expectancy ~ population, data = More)

# Results fit1
summary(fit1)

# Results fit2
summary(fit2)

```



## Conclusions

While the data does not show a statistically significant association between 
life expectancy and population size given our p-value is greater than our alpha 
level of 0.05 (p = 0.62), there was significant evidence of an association 
between life expectancy and infant mortality give out p-value is less than our 
alpha level of 0.05 (p < 0.001). Life expectancy decreases by 2.49 year on
average for every unit increase in infant mortality.


This section was added by Hayley Hemme

```{r}
world_2000 <- gapminder %>%
  filter(year == "2000")
```

Let's take a look at the distribution of fertility rates across the world in 2000. 
```{r, message=FALSE}
world_2000 %>% ggplot() + geom_histogram(aes(x=fertility)) + labs(title = "Histogram of Fertility Rates in 2000", x = "Fertility Rate")
```
There is a positively skewed distribution. Let's see if viewing fertility rates by continent will revel anything interesting.

```{r, message=FALSE}
world_2000 %>% ggplot() + geom_histogram(aes(x=fertility, fill=  continent)) + labs(title = "Histogram of Fertility Rates in 2000 by Continent", x = "Fertility Rate")
```
This plot reveals that highest average fertility rates occur in African countries, while lowest average fertility rates are seen in European countries.

Here is the same information plotted as a scatterplot with a fitted line for each continent.
```{r, message=FALSE}

world_2000 %>%  ggplot(aes(x = fertility, y = life_expectancy, color = continent)) + geom_point() + stat_smooth(se = F) + labs(title = "Fertility Rates and Life Expectancy in 2000 by Continent", x = "Fertility Rate", y = "Life Expectancy (years)")

```
Let's see how well fertility rate can predict life expectancy around the world in 2000.
```{r}
model1 <- (lm(life_expectancy ~ fertility, data = world_2000))
summary(model1)
```
This model seems to suggest that fertility is a good predictor of life expectancy. For every birth increase in average fertility rate, life expectancy decreases by 4.45 years (p < 0.0001). Let's see how this compares to infant mortality, which was found to be a statisitically significant predictor of life expectancy in Africa during the same year.

```{r}
model2 <- (lm(life_expectancy ~ infant_mortality, data = world_2000))
summary(model2)
```
Globally, there is a stronger correlation between infant mortality rate and life expectancy compared to fertility rate and life expectancy (R-squared = 0.79 vs. 0.64). For every unit increase in infant mortality rate, life expectancy decreases by 0.25 years (p < 0.0001).
