---
title: "dataanalysis_exercise"
format:
  html:
    theme: default
---

# Data Analysis Exercise (Module 4 Exercise)


## Description of Dataset

This data comes from the Centers for Disease Control and Prevention's National
Notifiable Diseases Surveillance System (NNDSS). Every week the CDC posts 
data received from the NNDSS system by the 50 states and US territories.
This data set in particular was posted on 01.26.2023. 



## Load and clean the data...
```{r}
# Load necessary packages...
library(tidyverse)

# Loading the raw data set.
RawNNDSS <- readr:: read_csv("data/NNDSS_Weekly_Data.csv")

## Let's take a look...
summary(RawNNDSS)
str(RawNNDSS)

## I'm kind of only interested in Georgia since I live here now...
GeorgiaNNDSS <- RawNNDSS %>%
  filter(`Reporting Area` == "GEORGIA")
summary(GeorgiaNNDSS)
str(GeorgiaNNDSS)

## What diseases DID Georgia from 2022 until now...
currentweek <- GeorgiaNNDSS %>%
  filter(`Current MMWR Year` >= 2022 & `Current week` != 'NA')

## Ehhhh let's get rid of 2022 for the sake of clear MMWR weeks...
currentweek <- GeorgiaNNDSS %>%
  filter(`Current MMWR Year` == 2022 & `Current week` != 'NA')

## Lets take a look at what diseases were MOST reported...
diseaseincidence <- currentweek %>%
  select(Label, `MMWR WEEK`,`Current week`) 

## Closer...
diseaseincidence1 <- currentweek %>%
  select(Label, `Current week`)

## Almost there...
diseaseincidence1 <- diseaseincidence1 %>%
  group_by(Label) %>%
  summarise(Freq = sum(`Current week`))
  
## Finally top ten have been narrowed down!
diseaseincidence1 <- diseaseincidence1 %>%
  arrange(desc(Freq)) %>%
  head(10)

```



## I'm a curious guy so I'm going to plot diseases reported over time and a 
## barplot to see the total burden of disease in Georgia in 2022...
```{r}
## You'll need two files for this...
saveRDS(diseaseincidence, file = "data/disease_incidence_time.rds")
saveRDS(diseaseincidence1, file = "data/disease_freq.rds")


# Load the thingy...
library(ggplot2)

# Disease incidence over time...
diseaseincidence %>%
  ggplot(mapping = aes(x = `MMWR WEEK`, y = `Current week`, color = Label)) +
  geom_line() + labs(x = "MMWR Week", y = "Incidence of Disease", title = 
                       "Disease Incidence Over Time (by Epi Week)")

# Wooooo that plot was extremely long, I'm going to remove the Legend to make
# it easier to see...
diseaseincidence %>%
  ggplot(mapping = aes(x = `MMWR WEEK`, y = `Current week`, color = Label)) +
  geom_line() + labs(x = "MMWR Week", y = "Incidence of Disease", title = 
                       "Disease Incidence Over Time (by Epi Week of 2022)") +
  theme(legend.position = "none")


# Disease Frequency in 2022
diseaseincidence1 %>%
ggplot(mapping = aes(x = reorder(Label, -Freq), y = Freq, fill = Label)) + 
  geom_bar(stat = 'identity') + 
  scale_x_discrete(labels = function(Label) str_wrap(Label, width = 6)) +
  theme(legend.position = "none") + geom_text(aes(label = Freq), vjust = 0) +
  labs(x = "Disease Frequency", y = "Disease Reported", 
       title = "Frequency of Top Ten Reported Disease in Georgia in NNDSS in 2022")

```

## Feel Free to start here...Have fun!

Everything from here on out was done by Katie Wells

Hi Connor! :) You already did some pretty cool analysis above, but I'm interested in the 10 least commonly reported infectious diseases in Georgia in this time period. Here we go!

```{r}
top_ten <- readRDS("data/disease_freq.rds")
all_ga <- readRDS("data/disease_incidence_time.rds")
```


```{r}
#make a new dataset with the 10 least reported diseases
least_incidence <- all_ga %>%
  group_by(Label) %>%
  summarise(Freq = sum(`Current week`)) %>% arrange(Freq) %>%
  head(10)

#let's take a look
print(least_incidence)
```
Cool! I'm glad to see that these diseases have a very low frequency in Georgia (but a little concerned that they might be hanging around). Let's put these in a bar graph to visualize it better. 

```{r}
#graphing the 10 least reported; I'll keep the same theme as above for cohesiveness
least_incidence %>%
ggplot(mapping = aes(x = reorder(Label, -Freq), y = Freq, fill = Label)) + 
  geom_bar(stat = 'identity') + 
  scale_x_discrete(labels = function(Label) str_wrap(Label, width = 6)) +
  theme(legend.position = "none") + geom_text(aes(label = Freq), vjust = 0) +
  labs(x = "Disease", y = "Cases Reported", 
       title = "Frequency of Ten Least Reported Diseases in Georgia in NNDSS in 2022")
```
Also cool! I wonder when these reports occurred in relation to MMWR weeks...

```{r}
#filter the all_ga data with the 10 least reported I found earlier
all_ga2 = filter(all_ga, Label %in% c("Haemophilus influenzae, invasive disease, Age <5 years, Unknown serotype", "Invasive pneumococcal disease, age <5 years, Confirmed", "Listeriosis, Confirmed", "Mumps", "Hepatitis, B, acute", "Tuberculosis", "Pertussis", "Haemophilus influenzae, invasive disease, All ages, all serotypes", "Shiga toxin-producing Escherichia coli (STEC)", "Varicella morbidity"))

#graph a scatterplot, see color legend above for particular disease
all_ga2 %>%
  ggplot(mapping = aes(x = `MMWR WEEK`, y = `Current week`, color = Label)) +
  geom_point() + labs(x = "MMWR Week", y = "Incidence of Disease", title = 
                       "Incidence of 10 Least Reported Diseases in Georgia Over Time") +
  theme(legend.position = "none")
```
Now we can see when these cases were reported based on MMWR week. It looks like these incidences were rather spread out across the year with only a couple of diseases having more than one case per week. 

All in all I though this was a really interesting dataset, and I like the way you processed the data as it led to some cool analyses!






