---
title: "Model Evaluation"
author: "Connor Ross"
editor: visual
---

# Model Evaluation

## Splitting the Data

```{r}
## Load packages
library(tidyverse)
library(tidymodels)
### Pathing(?)
library(here)
wrangled_data_location <- here("fluanalysis", "data", "processed_data", "SympActAny_new.rds")
SympActAny_new <- readRDS(wrangled_data_location)

## Splitting the wrangled dataset.
glimpse(SympActAny_new)

## For reproducibility
set.seed(1)

## Put 3/4 of the data into training set
data_split <- initial_split(SympActAny_new, prop = 3/4)

## Create data frame for two sets
train_data <- training(data_split)
test_data <- testing(data_split)

## New recipe
new_flu <- recipe(Nausea ~ ., data = train_data)

## Set engine/regression model
lr_mod <- logistic_reg() %>%
  set_engine("glm")


## Create workflow
flu_wflow <- workflow() %>%
  add_model(lr_mod) %>%
  add_recipe(new_flu)

flu_wflow


## Fit train data
flu_fit <- flu_wflow %>%
  fit(data = train_data)

flu_fit %>%
  extract_fit_parsnip() %>%
  tidy()


## Predict
predict(flu_fit, test_data)


## Augment
flu_aug <- augment(flu_fit, test_data)

flu_aug %>%
  select(Nausea, .pred_No, .pred_Yes)

## ROC curve
flu_aug %>%
  roc_curve(Nausea, .pred_No) %>%
  autoplot()


flu_aug %>%
  roc_auc(Nausea, .pred_No)





# RUNNY NOSE
new_flu_rn <- recipe(Nausea ~ RunnyNose, data = train_data)


## Create workflow
flu_wflow2 <- workflow() %>%
  add_model(lr_mod) %>%
  add_recipe(new_flu_rn)

flu_wflow2


## Fit train data
flu_fit_rn <- flu_wflow2 %>%
  fit(data = train_data)

flu_fit_rn %>%
  extract_fit_parsnip() %>%
  tidy()


## Predict
predict(flu_fit_rn, test_data)


## Augment
flu_aug_rn <- augment(flu_fit_rn, test_data)

flu_aug_rn %>%
  select(Nausea, .pred_No, .pred_Yes)

## ROC curve
flu_aug_rn %>%
  roc_curve(Nausea, .pred_No) %>%
  autoplot()


flu_aug_rn %>%
  roc_auc(Nausea, .pred_No)
```

# This Section Added by Kailin (Kai) Chen

## Set Seed for Reproducibility

```{r}
set.seed(325)
```

## Split Data Again

```{r}
# Data Splitting
split_data <- initial_split(SympActAny_new, prop = 4/5)

# Training & Test(ing) Data
train_data <- training(split_data)
test_data <- testing(split_data)
```

## Workflow Creation and Model Fitting

### Create Recipe for Fitting Linear Regression

```{r}
flu_recipe <- recipe(BodyTemp ~ ., data = SympActAny_new)
```

### Workflow to Create Linear Regression Model

```{r}
# Set Model to Linear Regression
linear_regression_model <- linear_reg() %>% set_engine("lm")

# Specifying Workflow
linear_workflow <- workflow() %>% 
  add_model(linear_regression_model) %>% 
  add_recipe(flu_recipe)

# Fitting/Training
linear_fit <- linear_workflow %>%
  fit(data = train_data)
```

## Multivariate Model Evaluation

### Prediction + RMSE

```{r}
# Fit Characteristics
linear_fit %>% extract_fit_parsnip() %>% tidy()

# Prediction
predict(linear_fit, test_data)

# Augmentation + RMSE
augment(linear_fit, test_data) %>% 
  select(BodyTemp, .pred) %>%
  rmse(BodyTemp, .pred)
```

## Simple (Alternative) Linear Regression Model Evaluation

### Edit Recipe

```{r}
simple_flu_recipe <- recipe(BodyTemp ~ RunnyNose, data = SympActAny_new)
```

### Modify Workflow

```{r}
# Modified Workflow
linear_workflow <- workflow() %>% 
  add_model(linear_regression_model) %>% 
  add_recipe(simple_flu_recipe)

# Fitting/Training
linear_fit <- linear_workflow %>%
  fit(data = train_data)
```

### Prediction + RMSE

```{r}
# Fit Characteristics
linear_fit %>% extract_fit_parsnip() %>% tidy()

# Prediction
predict(linear_fit, test_data)

# Augmentation + RMSE
augment(linear_fit, test_data) %>% select(BodyTemp, .pred) %>% rmse(BodyTemp, .pred)
```
